|

### 2022-07-23 manuscript.Rmd
still house-keeping

Although there are various ways to represent data patterns and models, visualization has been primarily taught in many data science courses for its efficiency. 
Such vision-dependent output may cause critical barriers against those who are blind and visually impaired, and sometimes people with learning disabilities. 
We argue that instructors need to teach multiple data representation methods so that all students of all abilities can benefit from multi-sensory data augmentation.

@@

Data visualization is the most common representation of data. 
Colors are an important part of accessibility of data visualizations. 
In our courses, while teaching colors in data visualizations, we introduce color-blindness through simulators. 
Those without any color vision deficiencies experience what data visualizations would look like to a sighted person that does have color blindness. 
Students get to experience this for red-, green-, and blue-blindness. 
For their data visualizations, instead of relying on the default color selection of R packages, students are introduced to the Okabe-Ito color palette which is known to be accessible to people with color vision deficiencies [@okabe]. 
In addition we rely on `scale_fill_viridis_d()` for generating `ggplot` visualizations that can be perceived by those with common forms of color blindness.

@@

# Closing Remarks
We firmly believe that an accessible future in the field of data science requires educators to deliberately teach accessibility as part of the curriculum. 
We also acknowledge that what we have presented here is only a small portion of the intersection of visual accessibility and data science. 
One of the co-authors was supported through a Teach Access grant to develop curricular materials on teaching accessibility as part of the data science curriculum. These open-access teaching materials can be accessed at [HIDDEN FOR REVIEW].

>>>>

Although there are various ways to represent data patterns and models, visualization has been primarily taught in many data science courses for its efficiency (cite). 
Such vision-dependent output may cause critical barriers against those who are blind and visually impaired, and sometimes people with learning disabilities. 
We argue that instructors need to teach multiple data representation methods so that all students of all abilities can benefit from multi-sensory data augmentation.

@@

Data visualization is the most common representation of data. 
Colors are an important part of accessibility of data visualizations. 
In our courses, while teaching colors in data visualizations, we introduce color-blindness through simulators. 
Those without any color vision deficiencies experience what data visualizations would look like to a sighted person that does have color blindness. 
Students get to experience this for red-, green-, and blue-blindness. 
For their data visualizations, instead of relying on the default color selection of R packages, students are introduced to the Okabe-Ito color palette which is known to be accessible to people with color vision deficiencies [@okabe]. 
In addition we rely on `ggplot2::scale_fill_viridis_d()` for generating `ggplot` visualizations [@ggplot22016] that can be perceived by those with common forms of color blindness.

@@

# Closing Remarks

We firmly believe that an accessible future in the field of data science requires educators to deliberately teach accessibility as part of the curriculum. 
We also acknowledge that what we have presented here is only a small portion of the intersection of visual accessibility and data science. 
One of the co-authors was supported through a Teach Access grant to develop curricular materials on teaching accessibility as part of the data science curriculum. These open-access teaching materials can be accessed at [HIDDEN FOR REVIEW].
